---
title: "How long does it take policy to be made?"
output: html_notebook
---

I want to collect congressional data to see if I can identify how long bills take to become law.
It will also be interesting to see what proportion of proposed bills become law.

```{r setup}
library(tidyverse)
library(rjson)
library(jsonlite)
library(RCurl)
library(lubridate)
library(tidyjson)
library(XML)
library(xml2)
library(tictoc)
library(furrr)
plan(multisession)
source("R/parsing_functions.R")

apiGovKey = getOption("apiGovKey")

```

# GovInfo API

[US General Publishing Office API documentation on github](https://github.com/usgpo/api)

API call format:

https:// api.govinfo.gov/published/`dateIssuedStartDate`/`dateIssuedEndDate`?offset=`startingRecord`&pageSize=`number of records in call`&collection=`comma-separated list of values`&api_key=`your api.data.gov api key`

Available collections:

```{r}
collection_url = URLencode(paste0("https://api.govinfo.gov/collections?api_key=", apiGovKey))

collections = flatten_dfc(fromJSON(collection_url))

glimpse(collections)
```

## Published

>This is similar to the collections service in that it provides users with an easy way to get a list of packages by date. The difference is that this service provides packages based on dateIssued -- this generally corresponds to the publication date of the content itself, rather than the govinfo system update time for a publication.

- Congressional Bills
- Congressional Bill Status



### Bill Status

Get a list of packages from the **Bill Status** collection which were issued in 2021.

```{r}

bill_status = getPublished(dateIssuedStartDate = ymd("2021-01-01"),
                           dateIssuedEndDate = ymd("2022-01-01"),
                           numRecords = 100,
                           collections = "BILLSTATUS")
glimpse(bill_status)
```

The return object is a list with a count of records, the API query for the next page, and a dataframe of packages. 
This dataframe contains:
- the package ID
- when it was issued and last modified
- the API query for the package
- the bill type
- the congress number

```{r}
glimpse(bill_status$packages)
```
Let's take a sample package and see what information we can get with that.

```{r}
sample_package = sample_n(bill_status$packages, 1)

sample_package
```

## Packages Service

>This service allows you to specify a govinfo `packageId` and retrieve available forms of content and metadata. A `/summary` json response is available that includes links and basic metadata about the package - generally equivalent to the information available on the details page for that package.
>
>From the summary, you can get access to all available content and metadata formats for a package.



```{r}
sample_package_summary = getPackages(packageId = sample_package$packageId, 
                                     summary = T) %>% 
  # If I want to create list column for download links
  # map_if(is_list, as_tibble) %>%
  # as_tibble()
  flatten_dfc()

sample_package_summary

```

### What's in the XML?

Now what can we get in the package's XML data

```{r}
(sample_xml = read_xml(sample_package_summary$xmlLink))

sample_xml_doc = xmlTreeParse(sample_xml, useInternalNodes = T)
top = xmlRoot(sample_xml_doc)

```



```{r}
sample_xml %>% 
  xml_child("bill")
```



```{r}
xml_find_all(sample_xml, "bill")
```



```{r}
(sample_bill = xml_child(sample_xml, "bill"))
```
So what will we want:

- billNumber
- createDate
- updateDate
- originChamber
- billType
- introducedDate
- congress
- recordedVotes
- committees
- committeeReports
- actions
- sponsors
- cosponsors

What happens with xmlToDataFrame?

```{r}
(sample_df = xmlToDataFrame(sample_xml_doc, nodes = getNodeSet(sample_xml_doc, "//bill")) %>% 
   select(billNumber, title, createDate:congress, policyArea)) %>% glimpse()
```

```{r}
sample_df$title
```


From this we can extract all of the attributes which only have a single value

```{r}
(xpaths = sample_bill %>% xml_children() %>% 
  map_chr(xml_path))
```

```{r}
(bill_nodesets = map(xpaths, ~ xml_find_all(sample_bill, xpath = .x)) %>% 
  set_names(map_chr(., xml_name)))
```
Some nodesets have multiple attributes so we'll need to work through those first

```{r}

(nested_attributes = bill_nodesets %>% 
  discard(~(xml_length(.) == 0)))

```

#### Committees

```{r}
# List of bill committees
(bill_committees = nested_attributes[["committees"]] %>% 
  xml_find_all("billCommittees/item"))
```

```{r}
bill_committees %>% 
  xml_structure()
```

```{r}
committees_list = bill_committees %>% 
  # retrieve each committee's xpath
  map_chr(xml_path) %>% 
  # Find the nodes based on each xpath
  map(~xml_find_all(sample_bill, .x)) %>% 
  # Coerce nodes to list
  map(as_list)

str(committees_list)
```
Because the list of committees is pretty deeply nested, we use `map_depth` and for each committee we use `modify_at` and `flatten_chr` to flatten out the list. 
For attributes which can have multiple values e.g. activities or subcommittees, we map over those and perform a similar procedure.
In each case we rename the variables according to their category and whittle them down to a dataframe.

```{r}
committees_dfs = committees_list %>% 
  map_depth(2, function(committee){
    committee %>% 
      modify_at("name", flatten_chr) %>% 
      modify_at("systemCode", flatten_chr) %>% 
      modify_at("chamber", flatten_chr) %>% 
      modify_at("type", flatten_chr) %>% 
      modify_at("activities", map, ~rename_with(flatten_dfc(.x), ~str_c("activity_", .))) %>% 
      modify_at("subcommittees", function(subcommittee){
          modify(subcommittee, function(item){
              modify_at(item, "systemCode", flatten_chr) %>% 
                  modify_at("name", flatten_chr) %>% 
                  modify_at("activities", map, ~rename_with(flatten_dfc(.x), ~str_c("activity_", .)))
          }) %>% 
              map(~rename_with(flatten_dfc(.x), ~str_c("subcommittee_", .)))
      }) %>% 
      flatten_dfc() %>% 
      rename_with(.fn = ~str_c("committee_", .), .cols = -starts_with("subcommittee")) %>% 
      select(starts_with("committee"), starts_with("subcommittee"))
  })
  
(committees_df = committees_dfs %>% 
  flatten_dfr())

```

#### Actions

```{r}

nested_attributes[["actions"]] %>% xml_path()
# List of bill committees
(bill_actions = nested_attributes[["actions"]] %>% 
  xml_find_all("item"))
```
```{r}
(actions_df = bill_actions %>% 
  # Coerce nodes to list
  map(as_list) %>% 
  map_dfr(function(action){
    action %>% 
      modify_at("sourceSystem", ~rename_with(flatten_dfc(.x), ~str_c("sourceSystem_", .))) %>% 
      modify_at("committees", function(committee){
          modify(committee, function(item){
              modify_at(item, "systemCode", flatten_chr) %>% 
                  modify_at("name", flatten_chr)
          }) %>% 
              map(~rename_with(flatten_dfc(.x), ~str_c("committee_", .)))
      }) %>% 
      flatten_dfc() %>% 
      rename_with(.fn = ~str_c("action_", .), .cols = -starts_with("action"))
  }))
```

#### Sponsors

```{r}
nested_attributes[["sponsors"]] %>% xml_path()
# List of bill sponsors
(bill_sponsors = nested_attributes[["sponsors"]] %>% 
  xml_find_all("item"))
```

```{r}
bill_sponsors %>% 
  xml_structure()
```

```{r}
(sponsors_df = bill_sponsors %>% 
  # Coerce nodes to list
  map(as_list) %>% 
  map_dfr(function(action){
    action %>% 
      modify_at("identifiers", ~rename_with(flatten_dfc(.x), ~str_c("identifiers_", .))) %>% 
      flatten_dfc() %>% 
      rename_with(.fn = ~str_c("sponsor_", .), .cols = -starts_with("sponsor"))
  }))
```

#### Cosponsors

```{r}
nested_attributes[["cosponsors"]] %>% xml_path()
# List of bill cosponsors
(bill_cosponsors = nested_attributes[["cosponsors"]] %>% 
  xml_find_all("item"))
```



```{r}
(cosponsors_df = bill_cosponsors %>% 
  # Coerce nodes to list
  map(as_list) %>% 
  map_dfr(function(cosponsor){
    cosponsor %>% 
      modify_at("identifiers", ~rename_with(flatten_dfc(.x), ~str_c("identifiers_", .))) %>% 
      flatten_dfc() %>% 
      rename_with(.fn = ~str_c("cosponsor_", .), .cols = -starts_with("cosponsor"))
  }))
```






```{r}
list(sample_df, committees_df, sponsors_df, actions_df)

(bill_combined_df = sample_df %>% 
  mutate(committees = list(committees_df), 
         sponsors = list(sponsors_df),
         actions = list(actions_df)))
```

### Putting this all together

With these data points parsed, I put together a number of functions (creatively named `parsing_functions` in the R folder) for processing an XML file.
Before I dive deeper into the API, I'm going to use the bulk data to explore first.

```{r}
fromJSON(URLencode(bill_status$nextPage))
```


## Bulk Data

Using the bulk data from https://www.govinfo.gov/bulkdata/BILLSTATUS/117/hr, we can 


```{r}
library(here)
hr117_files = list.files(here("data", "BILLSTATUS","117", "hr"), full.names = T)

(map_dfr(hr117_files[1:10], extract_bill_status) -> samples)
```

Using {furrr} has some benefits here

```{r}

tic("All hr117 files")
hr117_df = future_map_dfr(hr117_files, extract_bill_status)
toc()
```

### Explore HR117

Confirm that there is one bill number per row?

```{r}
add_count(hr117_df, billNumber, name = "rows_per_bill") %>% 
  count(rows_per_bill)
```

```{r}
glimpse(hr117_df)
```

```{r}
(hr117_counts = hr117_df %>% 
  mutate(# Count the number of rows in each element of each list column
         across(where(is_list), list(count = ~map_dbl(., nrow)), .names = "{.col}_{.fn}"),
         # Find whether the bill became law
         became_law = map_lgl(actions, ~any(str_detect(.$action_type, "BecameLaw")))) %>% 
  rowwise() %>% 
   # Get the date of the first and last actions
  mutate(date_first_action = min(actions$actionDate, na.rm=T),
         date_last_action = max(actions$actionDate, na.rm=T))) %>% 
  glimpse()
```

How often are each of the different elements found?

- How often do bills have cosponsors?
- How often are bills sent to committee?


```{r}
hr117_counts %>% 
  group_by(billType) %>% 
  # Count the number of times each status element is found
  summarise(bills = n(), 
            across(ends_with("_count"), ~sum(.>0, na.rm=T))) %>% 
  # Pivot and calculate this as a proportion
  pivot_longer(cols = c(bills:cosponsors_count), names_to = "element", values_to = "count", ) %>% 
  group_by(billType) %>% 
  mutate(pct = scales::percent( count/max(count) )) # Max shouldn't fail but could use exclude bills from pivot
```

How many committees are bills usually sent to?

```{r}

hr117_counts %>% 
  ggplot(., aes(x = committees_count))+
  geom_bar()+
  scale_x_continuous()+
  theme_minimal()
```

How many actions happen on a bill?

```{r}
hr117_counts %>% 
  ggplot(., aes(x = actions_count))+
  geom_bar()+
  scale_x_continuous()+
  theme_minimal()
```

 
```{r}
hr117_counts %>% 
  ggplot(., aes(x = cosponsors_count))+
  geom_bar()+
  scale_x_continuous()+
  theme_minimal()
```


 
```{r}
hr117_counts %>% 
  ggplot(., aes(x = votes_count))+
  geom_bar()+
  scale_x_continuous()+
  theme_minimal()
```

```{r}
hr117_counts %>% 
  select(billNumber, votes) %>% 
  unnest(votes) %>% 
  unnest(vote_roll) %>% 
  count(legislator, vote)
```

When are bills introduced?

```{r}
library(mokeR)

hr117_counts %>% 
  count(introducedDate, wt = n_distinct(billNumber)) %>% 
  group_by(week = week(introducedDate), year = year(introducedDate)) %>% 
  mutate(bills_weekly = sum(n),
         bill_week = mean(introducedDate)) %>% 
  ggplot(., aes(x = bill_week, y = bills_weekly))+
  geom_col()+
  scale_x_date(expand = expansion())+
  scale_y_continuous(expand = expansion())+
  theme_mark(plots_pane = T)
```

#### Actions

```{r}
hr117_counts %>% 
  select(billNumber, actions) %>% 
  unnest(actions) %>% 
  arrange(billNumber, actionDate, actionTime) %>% 
  select(billNumber, actionDate, actionTime, action_type, action_sourceSystem_name, everything())
```


```{r}
hr117_counts %>% 
  filter(date_first_action!=introducedDate) %>% 
  select(billNumber, introducedDate, date_first_action, actions) %>% 
  unnest(actions) %>% 
  filter(date_first_action == actionDate)
```



```{r}
hr117_actions = hr117_counts %>% 
  select(billNumber, actions) %>% 
  unnest(actions)

hr117_actions %>% 
  count(action_type)
```

```{r}
hr117_counts %>% 
  mutate(became_law = map_lgl(actions, ~any(str_detect(.$action_type, "BecameLaw")))) %>% 
  filter(became_law)
```


### History of Bills

```{r}
# History of bills modified since November 1st 2021
hob_2021 = getPublished(dateIssuedStartDate = ymd("2021-01-01"),
                           dateIssuedEndDate = ymd("2022-01-01"),
                           collections = "HOB")

hob_2021[["packages"]]
```

### Bill Status

```{r}
# History of bills modified since November 1st 2021
bill_status_2021 = getPublished(dateIssuedStartDate = ymd("2021-01-01"),
                           dateIssuedEndDate = ymd("2022-01-01"),
                           collections = "BILLSTATUS")

bill_status_2021[["packages"]]
```

Building the `getPackages` function

Aim: Return the history of bills data in a structure which can be analyzed

```{r}
bill_status_2021_pid = bill_status_2021[["packages"]][["packageId"]]
summary_url = paste0("https://api.govinfo.gov/packages/", bill_status_2021_pid, 
               "/summary",
               "?api_key=", apiGovKey)
summary_url

```

```{r}
# encode the URL with characters for each space.
summary_json = fromJSON(URLencode(summary_url[1]))

glimpse(summary_json)
```

```{r}
package_summary = flatten_dfc(summary_json)

glimpse(package_summary)

```

What is in the package MODS data?

```{r}
mods_url = URLencode(paste0(package_summary$xmlLink, "?api_key=", apiGovKey))

```



```{r}
xml_contents(package_mods)
```
```{r}
xml_children(package_mods) %>% glimpse()
```


```{r}
granules_url = paste0(package_summary$granulesLink, "&api_key=", apiGovKey)

granules_url_1000 = str_replace(granules_url, "pageSize=100", "pageSize=1000")
granules_url_1000

```

```{r}
granules = fromJSON(URLencode(granules_url))

granules

```

```{r}
granules_1000 = fromJSON(URLencode(granules_url_1000))

granules_1000

```


```{r}
granules_list = map(granules$granules$granuleLink, 
    ~fromJSON(URLencode(paste0(., "?api_key=", apiGovKey))))

granules_list[[7]][["download"]][["txtLink"]] %>% 
  str_c(., "?api_key=", apiGovKey) %>% 
  read_html()
```

```{r}
granules_flattened = granules_list %>% 
    map_dfr(discard, is_list)

glimpse(granules_flattened)

```
```{r}
granules_tables = granules_list %>% 
    map(keep, is_list)

glimpse(granules_tables[1:3])
```
```{r}
granules_df = granules_flattened %>% 
  mutate(download = map(granules_tables, "download"),
         committees = map(granules_tables, "committees"),
         members = map(granules_tables, "members"),
         references = map(granules_tables, "references")) %>% 
  unnest(committees, keep_empty = T, names_sep = "_") %>% 
  unnest(references, keep_empty = T, names_sep = "_") %>% 
  unnest(references_contents, keep_empty=T, names_sep = "_")

glimpse(granules_df)
```
```{r}
granules_df %>% 
  add_count(granuleId) %>% 
  filter(n>1)
```

Get XML from granules

```{r}
granules_xml_links = granules_list %>% 
  map("download") %>% 
  map_dfr(flatten_dfr) %>% 
  head(1) %>% 
  transmute(modsLink = str_c(modsLink, "?api_key=", apiGovKey)) %>% 
  pull()

granules_xml = granules_xml_links %>% 
  map(read_xml)

granules_result = granules_xml[[1]] %>% xmlParse()

granules_result
```
```{r}
as_list(granules_xml[[1]])
```



Get package summary data

```{r}
getPackages(hob_2021[["packages"]][["packageId"]], summary=T) %>% 
  flatten_dfc()
```

Get package granules data

```{r}
hob_2021_granules = getPackages(hob_2021[["packages"]][["packageId"]], granules=T)

hob_2021_granules
```

```{r}
hob_2021_granules$committees[1:2]
  map("contents") %>% 
  flatten_dfr()
```



# Resources

- https://www.earthdatascience.org/courses/earth-analytics/get-data-using-apis/API-data-access-r/
- https://stackoverflow.com/questions/43843204/r-error-when-using-geturl-from-curl-after-site-was-changed
- https://robotwealth.com/how-to-wrangle-json-data-in-r-with-jsonlite-purr-and-dplyr/
