---
title: "Explore Bill Status bulk data"
output: 
  rmdformats::robobook:
    df_print: paged
editor_options: 
  markdown: 
    wrap: sentence
---

The goal of exploring legislative data is to provide a clearer view of who our representatives are and what they are doing.
An ideal solution combines Datamade's [beautifully simple representative finder](https://myreps.datamade.us/) and Project Vote Smart's [vast collection of data on legislator activities](https://justfacts.votesmart.org/).

Other questions worth investigating:

-   how long does it take for a bill to become law?
-   which legislative processes are the most time-intensive?
-   which decision-makers are acting in each stage?
-   what legislative paths can a bill take? map the legislative process for bills and get a sense for how legislator behaviour affects a bill's likelihood of being considered, its content, and the eventual outcome
-   Make a sankey diagram depicting how bills flow through congress

It might be worth reaching out to: <https://github.com/unitedstates/congress>

```{r setup}
library(extrafont)
library(tictoc)
library(furrr)
library(here)
library(reactable)
plan(multisession)
source("R/parsing_functions.R")

action_codes = read_csv(here("data", "action_codes.csv"), col_types = "cc")
```

# Collecting the data

I have downloaded all the different bill types for the 117th congress from <https://www.govinfo.gov/bulkdata/BILLSTATUS/117/>.
Below is the structure of this directory

```{r dir-tree}
fs::dir_tree(here("data", "BILLSTATUS", "117"), recurse = F)
```

First we'll parse bills which were introduced in the house, or the bill types beginning with "h".
With the directory structure, I create a list of files by bill type.

House-originated bills (H.R.) are by far the most common, while simple resolutions (H. Res.) are the second most common.
Joint resolutions (H.J. Res.) and concurrent resolutions are introduced much less frequently.

```{r list-bill-files}
house_bill_types = list.files(here("data", "BILLSTATUS","117"), pattern = "^h")

(house_bill_folders = here("data", "BILLSTATUS", "117", 
                          house_bill_types) %>% 
    str_remove_all("\\/OneDrive"))

house_bill_files = map(house_bill_folders, list.files, full.names = T) %>% 
  set_names(house_bill_types)

```

How many bill XML files are in each folder?

```{r count-bills}
house_bill_files  %>% 
  map(length)
```

# Examine a single bill

To start off, I want to parse a single bill.

```{r sample-bill}
# Sample a bill
sample_bill_file = sample(house_bill_files$hr, 1)

# Extract the bill status data
(sample_bill_df = extract_bill_status(sample_bill_file, 
                                     get_votes = F,
                                     log_types = "console"))

# View the bill's contents
glimpse(sample_bill_df)

```

Let's put this into a a nicer table (Note I have a preference for the date format `dd/mm/yyyy` so that is the format used here)

```{r sample-bill-rt}
select(sample_bill_df, 
       bill_id, title, introducedDate, where(negate(is_list)), createDate,
       -constitutionalAuthorityStatementText) %>% 
  reactable(
    columns = list(
      originChamber = colDef(name = "Chamber of Origin", width = 80),
      bill_id = colDef(name = "ID", width = 70),
      congress = colDef(show = F, width = 60),
      billType = colDef(show = F,name = "Bill Type", width = 50),
      billNumber = colDef(show = F,name = "Bill #", width = 40),
      title = colDef(name = "Bill Title", width = 420),
      createDate = colDef(name = "Created", width = 80,
                          format = colFormat(date = T, locales = "en-GB")),
      latestAction_actionDate = colDef(name = "Date", width = 80,
                                       format = colFormat(date = T, locales = "en-GB")),
      latestAction_text = colDef(name = "Action", width = 320),
      version = colDef(show=F),
      updateDate = colDef(name = "Updated", width = 80,
                          format = colFormat(date = T, locales = "en-GB")),
      introducedDate = colDef(name = "Introduced", width = 80, 
                              format = colFormat(date = T, locales = "en-GB"))
    ),
    columnGroups = list(
      colGroup(name = "Latest Action",
               columns = c("latestAction_actionDate", "latestAction_text"))
    )
  )
```

## Actions

What actions are taken on a bill?
After a bill is introduced, how do members of congress interact with it?

To answer questions like these, we need to `unnest()` the `actions` column.

```{r sample-actions}
(sample_actions = sample_bill_df %>% 
  select(billType, billNumber, bill_id, title, createDate, updateDate, title,
         -where(is_list), actions) %>%
  unnest(actions) %>% 
   select(bill_id, 
          action_type, actionCode, action_text, any_of(c("actionDate", "actionTime")), 
         action_source_name, action_source_code,
         action_committee_name, action_committee_systemCode,
         everything()) %>% 
   # Order bills and actions
  arrange(billType, billNumber, actionDate))
```

Let's visualize what this timeline might look like, counting the number of each type of action taken.

```{r sample-action-timeline}
sample_actions %>% 
  # Sort cols and rows
  select(bill_id, billType, billNumber, action_type, actionDate, action_text, action_committee_name, action_source_name) %>%
  mutate(action_type = fct_explicit_na(action_type, na_level = "(Null Action Type)")) %>% 
  arrange(billType, billNumber, actionDate, action_type) %>% 
  # Plot
  ggplot(., aes(y = action_type, x = actionDate))+
  # Line connecting points
  geom_line(aes(group = bill_id), size = .8, alpha = .9, colour = my_col_pal[2])+
  # Points using the count of each action type on a given date
  geom_jitter(aes(colour = action_type), 
              size = 3, alpha = .85, 
              show.legend = F,
              height = 0.025, width = 2.5)+
  labs(title = "Sample bill timeline",
       subtitle = str_c("Timeline of actions for bill ", sample_bill_df$bill_id, ", by action type"))+
  scale_x_date(name = "Action Date", breaks = scales::breaks_pretty(n = 6), labels = scales::label_date_short())+
  # scale_y_continuous(name = "# of actions", 
  #                    breaks = scales::breaks_width(1), limits = c(0, NA))+
  scale_colour_manual(values = viz_colours[1:length(unique(sample_actions$action_type))])+
  scale_size(guide = "none", range = c(2,7))+
  theme_moke(plots_pane = T)+
  theme(legend.position = "top", legend.justification = "left",
        axis.title.y.left = element_blank(), axis.title.x = element_blank(),
        axis.ticks.x = element_line())
```

However, it looks like we will need to clean the actions because there appears to be a lot of duplicate `IntroReferral` actions.
Below is a table of these actions in our sample bill: `r sample_bill_df$bill_id`

```{r}
sample_actions %>% 
  filter(action_type == "IntroReferral") %>% 
  select(bill_id, title, actionDate, action_type, actionCode, action_text, action_source_name, action_committee_name)
```


## Create House bills dataset

```{r sample-hr-bills, eval=F}
# Test for sample
# Extract bill data for a sample of house bills
hr_sample = sample(house_bill_files$hr, 1000)
```

```{r extract-sample-bills, eval=F}
tic(str_c("Extract ", length(hr_sample), " bills"))
(hr_bills = hr_sample %>% # house_bill_files$hr %>% 
  future_map_dfr(extract_bill_status, get_votes = F, log_types = "console") %>% 
    mutate(
      across(ends_with("Date"), as_datetime),
      actions = map(actions, mutate, 
                    action_type = factor(action_type, 
                                         levels = c("IntroReferral", "Committee", "Floor", 
                                                    "Discharge", "President", "BecameLaw"), 
                                         ordered = T))
      ))
toc()

```

```{r sample-hr-glimpse, eval=F}
glimpse(hr_bills)
```

First combine the bill file list into a single vector and question whether they should even be in a list.
Then run our `extract_bill_status()` function on all of our bills, in parallel using `{furrr}` to save some time.

```{r, eval=F, echo = T}
all_files = flatten_chr(house_bill_files)

tic(str_c("Extract ", sum(map_dbl(house_bill_files, length)), " bills"))
(all_bills = future_map_dfr(all_files, extract_bill_status, log_types = "console"))
toc()
```

Save this object for use later

```{r}
# saveRDS(all_bills, here("data", "cleaned", "BILLSTATUS_117_House.Rds"))
all_bills = readRDS(here("data", "cleaned", "BILLSTATUS_117_House.Rds"))
```

## Parse Actions

The actions tibbles are not necessarily structurally consistent.
What fields are present and when?

```{r}
all_bills %>% 
  select(bill_id, actions) %>% 
  # Create a list of field names for each actions dataframe
  mutate(actions_cols = map(actions, colnames)) %>% 
  unnest(actions_cols) %>%
  # Count the number of bills the field is available for
  count(actions_cols, wt = n_distinct(bill_id), sort = T)
```

Committee information is missing for a very small number of cases and timing information is missing for the vast majority of actions.
When is time information available?

```{r}
all_bills %>% 
  select(bill_id, actions) %>%
  unnest(actions) %>% 
  # Filter to rows with a time
  filter(!is.na(actionTime)) %>% 
  count(action_type, sort = T)
```

Now, let's `unnest()` the actions data and see what action-level information looks like.

```{r}
(actions_unnested = select(all_bills, where(negate(is_list)), starts_with("action")) %>% 
   # Unnest actions
  unnest(actions) %>% 
   # Remove duplicate actions
   # - Intro-H
   filter(actionCode != "Intro-H"|is.na(actionCode)) %>% 
   # Create action timestamp
  mutate(action_ts = make_datetime(year = year(actionDate), 
                                   month = month(actionDate), day = day(actionDate), 
                                   hour = coalesce(hour(actionTime), 0), 
                                   min = coalesce(minute(actionTime), 0), 
                                   sec = coalesce(second(actionTime), 0), 
                                   tz = "US/Eastern"),
         action_type = factor(action_type, 
                              levels = c("IntroReferral", "Committee", "Floor", 
                                         "Discharge", "President", "BecameLaw"), 
                              ordered = T),
         action_source_name = factor(action_source_name,
                                     levels = c("Library of Congress", "House floor actions", 
                                                "House committee actions", "Senate"),
                                     ordered = T)) %>% 
   
   # Join in action codes
   left_join(action_codes, by = c("actionCode" = "Code")) %>% 
   # Order bills and actions
   arrange(bill_id, action_ts, action_type, action_source_name) %>% 
   # Number actions
  group_by(billType, billNumber) %>% 
   # Number actions
  mutate(action_number = row_number(),
         became_law = ("BecameLaw" %in% action_type), .after = actionTime) %>% 
  ungroup())
```


What action types are most common?

```{r}
actions_unnested %>%
  count(billType, action_type) %>% 
  arrange(billType, desc(n)) %>% 
  pivot_wider(names_from = action_type, values_from = n, values_fill = 0)
```

Or in a graph:

```{r}
actions_unnested %>%
  count(billType, action_type) %>% 
  arrange(billType, desc(n)) %>% 
  ggplot(., aes(y = billType, fill = action_type))+
  geom_col(aes(x = n), position = position_fill())+
  scale_x_continuous(name = NULL, labels = scales::percent, breaks = scales::breaks_pretty())+
  scale_fill_manual(name = "Action", 
                    values = viz_colours[1:length(unique(actions_unnested$action_type))])+
  # facet_wrap(~billType, scales = "free_y")+
  theme_moke(plots_pane = T)
```

```{r}
actions_unnested %>% 
  group_by(became_law, billType, bill_id, action_type, actionDate) %>% 
  summarise(actions = n()) %>% 
  ggplot(., aes(x = actionDate, y = actions))+
  geom_point(aes(colour = action_type))+
  facet_wrap(~became_law)+
  theme_moke(plots_pane = F)
```



#### Action Codes

What about action codes?

```{r}
actions_unnested %>% 
  select(bill_id, actionDate, actionTime, action_type, actionCode, Action) %>% 
  count(actionCode, Action, sort = T)
```

```{r}
actions_unnested %>% 
  group_by(bill_id) %>%
  summarise(
    actions = n(),
    actions_w_code = sum(!is.na(Action)),
    actions_coded = actions_w_code/actions
  ) %>% 
  ggplot(., aes(x = actions_coded))+
  geom_histogram()+
  scale_x_continuous(labels = scales::percent)+
  scale_y_continuous(labels = scales::comma, expand = expansion())+
  theme_moke(plots_pane = T)
```

The actions data is messy.
There are a number of cases where actions are duplicated in the data which need to reconcile in order to make sense of the legislative timeline.
Below are a list of the cases I have found:

- Action code `E30000` and `36000` with action type `President` and `BecameLaw` respectively appear to both represent the president signing a bill into law, while action codes `E40000` and `36000` with the same respective action types represent a bill becoming a public law. In this case, action type `President` and `BecameLaw` are interchangeable
- Action code `Intro-H` appears to be a duplicate of action code `1000`, both of which are `IntroReferral` action types


How do actions with actionCodes listed on congress.gov differ from those which are not?

```{r}
actions_unnested %>% 
  filter(bill_id == sample(bill_id, 1)) %>% 
  select(bill_id, actionDate, actionTime, action_type, actionCode, Action, action_text, starts_with("action"))
```


How do action codes relate to action sources?

```{r}
actions_unnested %>% 
  count(action_source_name, actionCode, str_trunc(action_text, 50), sort = T)
```

Is there a difference between the `actionCode` **1000** and **Intro-H**?

```{r}
actions_to_compare = c("1000", "Intro-H")

actions_unnested %>% 
  filter(actionCode %in% actions_to_compare)
  
```

Do these two actions occur for every bill?

```{r}
actions_unnested %>% 
  filter(actionCode %in% actions_to_compare) %>% 
  count(actionCode)
```

No, **1000** occurs more often.

Does `Intro-H` **only occur** in cases where `1000` is present too?

```{r}
actions_unnested %>% 
  group_by(bill_id) %>% 
  filter(actionCode %in% actions_to_compare,
         "Intro-H" %in% actionCode) %>% 
  ungroup() %>% 
  count(actionCode)
```

So effectively we can remove `Intro-H` actions.
This reduces the size of the actions data by ~8,000 rows.

```{r}
actions_unnested %>% 
  select(bill_id, title, introducedDate, 
         actionDate, actionTime, action_type, actionCode, action_text,
         starts_with("action_source"), starts_with("action_committee")) %>% 
  
  filter(actionCode != "Intro-H"|is.na(actionCode)) %>% 
  mutate(action_ts = make_datetime(year = year(actionDate), month = month(actionDate), day = day(actionDate), 
                                   hour = coalesce(hour(actionTime), 0), 
                                   min = coalesce(minute(actionTime), 0), 
                                   sec = coalesce(second(actionTime), 0), 
                                   tz = "US/Eastern"), .before = actionDate) %>% 
   arrange(bill_id, action_ts)
```



```{r, eval=F}
# Reduce data down to bill-level attributes and the actions data
(house_bill_actions = select(all_bills, billType, introducedDate, policy_areas, where(negate(is_list)), starts_with("action")) %>% 
   # Unnest actions
  unnest(actions) %>% 
   # Remove duplicate actions
   # - Intro-H
   filter(actionCode != "Intro-H"|is.na(actionCode)) %>% 
   # Create action timestamp
  mutate(action_ts = make_datetime(year = year(actionDate), month = month(actionDate), day = day(actionDate), 
                                   hour = coalesce(hour(actionTime), 0), 
                                   min = coalesce(minute(actionTime), 0), 
                                   sec = coalesce(second(actionTime), 0), 
                                   tz = "US/Eastern"),
         action_type = factor(action_type, 
                              levels = c("IntroReferral", "Committee", "Floor", 
                                         "Discharge", "President", "BecameLaw"), 
                              ordered = T),
         action_source_name = factor(action_source_name,
                                     levels = c("Library of Congress", "House floor actions", 
                                                "House committee actions", "Senate"),
                                     ordered = T)) %>% 
   
   # Join in action codes
   left_join(action_codes, by = c("actionCode" = "Code")) %>% 
   # Order bills and actions
   arrange(bill_id, action_ts, action_type, action_source_name) %>% 
   # Number actions
  group_by(billType, billNumber) %>% 
   # Number actions
  mutate(action_number = row_number(),
         became_law = ("BecameLaw" %in% action_type), .after = actionTime) %>% 
  ungroup()) %>% 
  glimpse()
```

### Policy areas

What policy areas are present:

```{r}
(policy_areas = all_bills %>% 
  unnest(policy_areas, keep_empty = T)) %>% 
  count(policy_areas, sort = T)
```

What about the different legislative subjects?

```{r}
(legislative_subjects = policy_areas %>% 
  unnest(legislative_subjects, keep_empty = T)) %>% 
  count(legislative_subjects, sort = T)
```
Are legislative subjects missing for a certain bill type?

```{r}
legislative_subjects %>% 
  add_count(billType, name = "total_bills", wt = n_distinct(bill_id)) %>% 
  filter(!is.na(legislative_subjects)) %>% 
  count(billType, total_bills, wt = n_distinct(bill_id)) %>% 
  mutate(have_leg_subject = n/total_bills)
```


```{r}
legislative_subjects %>% 
  add_count(policy_areas, name = "policy_area_count", wt = n_distinct(bill_id)) %>% 
  count(policy_areas, legislative_subjects, policy_area_count) %>% 
  arrange(-policy_area_count, -n)
```

Look at Government Operations and Politics legislation

```{r}
(gov_ops_bills = legislative_subjects %>% 
  # filter("Government Operations and Politics" %in% policy_areas)) # if single row per bill
   filter(policy_areas == "Government Operations and Politics"))
```

```{r}
gov_ops_bills %>% 
  select(bill_id, ends_with("Date"), legislative_subjects) %>% 
  add_count(legislative_subjects, wt = n_distinct(bill_id)) %>% 
  filter(n > 100) %>% 
  pivot_longer(
    cols = ends_with("Date"), 
    names_to = "type", 
    values_to = "time", 
    names_transform = list(type = ~str_replace_all(., c("Date" = "", "(?=_).+" = "")))) %>% 
  mutate(date = as.Date(time)) %>% 
  ggplot(., aes(x = time, y = bill_id))+
  # Line for each bill, point for each date
  geom_line(size = 1)+
  geom_point(aes(colour = type), size = 3.5, alpha = .8)+
  # Axes
  scale_x_datetime(name = NULL, labels = scales::label_date_short())+
  scale_y_discrete(name = NULL, position = "right")+
  scale_colour_manual(name = "Date", values = viz_colours[1:4])+
  # Facet by legislative subject
  facet_wrap(~legislative_subjects, scales = "free_y", ncol = 1, 
             strip.position = "left")+
  # Theming
  theme_moke(plots_pane = T, text_family = NULL, title_family = NULL, subtitle_family = NULL)+
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(colour = my_col_pal[4]),
        axis.line = element_blank(),
        legend.position = "top")
```

Let's look at the Administrative law and regulatory procedures bills

```{r}
(admin_law_bills = gov_ops_bills %>% 
  filter(legislative_subjects == "Administrative law and regulatory procedures") %>% 
  select(1:2, title, latestAction_text, latestAction_actionDate, everything()))
```

```{r}
admin_law_bills %>% 
  select(bill_id, title, actions) %>% 
  unnest(actions) %>% 
  mutate(action_type = factor(action_type, 
                              levels = c("IntroReferral", "Committee", "Floor", 
                                         "Discharge", "President", "BecameLaw"), 
                              ordered = T),
         action_source_name = factor(action_source_name,
                                     levels = c("Library of Congress", "House floor actions", 
                                                "House committee actions", "Senate"),
                                     ordered = T)) %>% 
  # count(action_type)
  select(bill_id, title, 
         action_text, action_type, actionCode, actionDate, actionTime, 
         starts_with("action_source"), starts_with("action_committee")) %>% 
  # Remove intro actions with a missing action code
  filter(actionCode != "Intro-H"|is.na(actionCode)) %>%
  # Create action timestamp to incorporate both date and time
  mutate(action_ts = make_datetime(year = year(actionDate), 
                                   month = month(actionDate), 
                                   day = day(actionDate), 
                                   hour = coalesce(hour(actionTime), 0), 
                                   min = coalesce(minute(actionTime), 0), 
                                   sec = coalesce(second(actionTime), 0), 
                                   tz = "US/Eastern"), 
         .before = actionDate) %>% 
   arrange(bill_id, action_ts, action_type, action_source_name) %>% 
   # Number actions
  group_by(bill_id) %>% 
  mutate(action_number = row_number(),
         became_law = ("BecameLaw" %in% action_type), .after = actionTime) %>% 
  ungroup() %>% 
  select(-action_ts, -became_law) %>% 
  reactable(resizable = T)
```

When are these bills created/updated?

```{r}
all_bills %>% 
  # Calculate summary stats for each date var
  summarise(
    across(c(createDate, updateDate),
           .fns = list(min = min, mean = mean, max = max, median = median), 
           na.rm=T, .names = "{.fn}_{.col}")
  ) %>% 
  # pivot for table
  pivot_longer(
    everything(),
    names_to = c("type", ".value"),
    names_sep = "_",
    values_to = "stat"
    )
```

What does this look like on a histogram?

```{r}
all_bills %>% 
  select(bill_id, billType, ends_with("Date")) %>% 
  pivot_longer(
    cols = ends_with("Date"), 
    names_to = "type", 
    values_to = "time", 
    names_transform = list(type = ~str_replace_all(., c("Date" = "", "(?=_).+" = "")))) %>% 
  mutate(date = as.Date(time)) %>% 
  ggplot(., aes(x = date, fill = type))+
  geom_histogram(alpha = .8,
                 binwidth = 14, position = position_identity())+
  facet_grid(billType~type, scales = "free_y")+
  theme_moke(plots_pane = T, title_family = NULL, subtitle_family = NULL, text_family = NULL)
```










View number of actions taken on a bill by whether or not they became law

```{r fig.height = 7, fig.width = 8}
actions_unnested %>% 
  group_by(bill_id) %>% 
  mutate(became_law = ("BecameLaw" %in% action_type)) %>% 
  ggplot(., aes(x = action_ts, y = action_number))+
  geom_line(aes(group = bill_id, colour = became_law, alpha = became_law, size = became_law))+
  scale_x_datetime(name = NULL, expand = expansion())+
  scale_y_continuous(expand = expansion())+
  scale_alpha_manual(values = c(.5, .9))+
  scale_size_manual(values = c(.4, 1))+
  scale_colour_manual(values = my_col_pal[c(3,6)])+
  labs(title = "Cumulative bill actions, by legislation type and whether it became law")+
  facet_wrap(~billType, scales = "free_y")+
  theme_moke(plots_pane = T)
```


Can we get an idea of the order of actions based on their average position in the data?

```{r}
(action_order_summary = actions_unnested %>% 
  group_by(billType, action_type) %>% 
  summarise(across(action_number, list(min = min, 
                                       y25 = ~quantile(., .25), 
                                       mean = mean, median = median, 
                                       y75 = ~quantile(., .75),
                                       max = max), na.rm=T)))
```

```{r}
action_order_summary %>%
  pivot_longer(cols = starts_with("action_n"), names_to = "measure", values_to = "value",
               names_transform = list(measure = ~str_remove_all(., "^action_number_")))

action_order_summary %>% 
  ggplot(., aes(x = action_type))+
  geom_boxplot(
    aes(ymin = action_number_min, 
        lower = action_number_y25, middle = action_number_median, upper = action_number_y75,
        ymax = action_number_max),
    stat = "identity"
  )+
  facet_wrap(~billType)+
  theme_moke(plots_pane = T)+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r}
actions_unnested %>% 
  group_by(action_type) %>% 
  mutate(mean_num = mean(action_number)) %>% 
  ungroup() %>% 
  arrange(mean_num) %>% 
  mutate(action_type = fct_inorder(factor(action_type))) %>% 
  ggplot(., aes(x = action_type, y = action_number))+
  geom_boxplot()+
  labs(title = "In what order do actions typically occur?", x = NULL)+
  facet_wrap(~billType)+
  theme_moke(plots_pane = T)

```

```{r}
actions_unnested %>% 
  filter(bill_id == sample(bill_id, 1)) -> sample_bill_actions

sample_bill_actions %>% 
  # Plot
  ggplot(., aes(x = actionDate))+
  # Line connecting points
  geom_line(aes(y = ..count.., group = bill_id), stat = "count", size = .8, alpha = .8)+
  # Points using the count of each action type on a given date
  geom_point(aes(colour = action_type, size = ..count..), stat = "count")+
  labs(title = "Bill actions over time", 
       x = "Action Date", y = "# of actions")+
  scale_size(guide = "none", range = c(2,7))+
  theme_moke(plots_pane = T)
```




Let's look only at those bills that became law

```{r}
actions_unnested %>% 
  # Combine bill type and number to create an ID
  unite(bill_id, billType, billNumber, sep = "-", remove = F) %>% 
  group_by(bill_id) %>% 
  filter("BecameLaw" %in% action_type) %>%
  ungroup() %>% 
  ggplot(., aes(x = action_ts, y = action_number))+
  geom_line(aes(group = bill_id))+
  facet_wrap(~billType, scales = "free_y")
```


```{r}
actions_unnested %>% 
  # Combine bill type and number to create an ID
  unite(bill_id, billType, billNumber, sep = "-", remove = F) %>% 
  group_by(bill_id) %>% 
  filter("BecameLaw" %in% action_type)
```




```{r}
sample_bill_df[["actions"]][[1]] %>% 
  distinct(action_type, actionDate, action_committee_name)
```


Why are there multiple of what seem to be the same action?

```{r}
read_xml(sample_bill_file) %>% 
  xml_child("bill") %>% 
  xml_child("actions") %>% 
  xml_find_all("item") %>% 
  map(xml_contents)
```


